Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-11-30 inst/paper/main.tex

\begin{lstlisting}
  > library("ggplot2")  # load a library from string
  > library(ggplot2)     # equivalent, except symbol is not in scope

  library <- function(package, ...) {
    ...
    if(!is.character(package))
      package <- as.character(substitute(package))
    ...
  }
\end{lstlisting}

@@

\section{Evaluation}

This section presents the results of our empirical study of \eval in R
in terms of four research questions.

\subsection{Use of Eval}

The first research question we address is how and why \eval is used in R code.

\begin{center}{\bfseries RQ1: What are the patterns of usage of \eval in the corpus?}\end{center}

This section presents an overview of the usage of \eval in our corpus. Richards
et al. reported 82\% of JavaScript-enabled web pages using \eval~\cite{ecoop11}.
The case for R is more striking with 100\% of the programs call the \eval
function.\footnote{We combine calls to {\sf eval}, {\sf eval.parent}, {\sf
    evalq} and {\sf local}.} This high number is explained by the fact that the
R implementation, its core packages, use \eval extensively. We differentiate
three types of eval based on their origin:
%
\begin{compactitem}[$-$]
  \item a {\bf core eval} is one that occurs in the code of a core package,
  \item a {\bf dependent eval} occurs in a package imported by the target
  package, and
  \item an {\bf own eval} is one that occurs in the source code of the target
  package.
\end{compactitem}

Over all, we observe \AllAllCallCountRnd calls to \eval in our corpus.
Table~\ref{A} details this run-time information. The format is T (m/M/x)
where T is the total number of calls observed for all runs of the corpus, m
and x are the minimum and maximum number of calls for any given run, and M
is the mean number of calls per run. An average CRAN package run triggers
581 {\eval} calls in core package, 532 in dependent packages and 292 in the
CRAN package itself.  The Kaggle data set has small programs which only
trigger \eval in core and dependent packages. Thus, an average Kaggle run
has 412 core calls, 96 dependent calls, and no own calls.

%% SORRY  I HAD TO REFORMAT
\begin{table}[!h] \centering
 \begin{tabular}{r|c|c|c} \hline
Source & Core    & Dependent                      & Own \\\hline
CRAN   & 5.6M &  867K    & 259K \\
       & (1 / 337 / 46K)& (0 / 52 / 58K ) & (0 / 15 / 15K) \\\hline
Kaggle &  1.2M &  470K   & 0\\
 &  (0 / 721 / 18K) & (0 / 290 / 20K) & (0 / 0 / 0)\\\hline
\end{tabular}\caption{Eval calls in CRAN and Kaggle}\label{A}\end{table}

Table~\ref{table:site-package-summary} shows the number of call sites in
CRAN packages. \OneCallSitesEnrich{} packages have a single call site, and one
package, \emph{metafor}, has \MaxCallSitesEnrich{} call sites of
\eval. \NoCallSitesEnrich{} packages have no calls to \eval. The core packages
have \NbCoreEvalCallSites{} call sites of \eval and Kaggle has none. These
numbers are under-approximations as \eval may be called reflectively,
through aliases, or passed to a higher-order function.

\begin{table}[ht]\resizebox{\columnwidth}{!}{%
\begin{tabular}{rr|rr|rr}\hline
 Callsites & Packages & Callsites & Packages & Callsites & Packages \\\hline
\input{tag/table-site-package-summary.tex}
\end{tabular}}
\caption{Distribution of callsites in CRAN}
\label{table:site-package-summary}
\end{table}

We compare the number of events intercepted during execution of \eval across
the whole corpus. For this, we use the number of interpreter steps as a
proxy for events. We observe \AllEventCountRnd events all told, and
\EvalEventCountRnd events occur inside eval (\EvalEventAllPerc of all
events). Table~\ref{table:event-distribution} shows the distribution of
events caused by \eval calls.

\begin{table}[ht]\resizebox{\columnwidth}{!}{%
\begin{tabular}{lll}\hline
Events & Core Eval Call \% & Package Eval Call \% \\ \hline
\EventsMinRangeA--\EventsMaxRangeA & \EventsCoreEvalPercA & \EventsPackageEvalPercA \\
\EventsMinRangeB--\EventsMaxRangeB & \EventsCoreEvalPercB & \EventsPackageEvalPercB \\
\EventsMinRangeC+ & \EventsCoreEvalPercC & \EventsPackageEvalPercC \\ \hline
\end{tabular}} \label{table:event-distribution}
  \caption{Events in \eval calls}
\end{table}

@@

% Following the high-level overview of the usage of eval in the previous section,
% we now look into categorization of the eval behavior. Similarly to the JavaScript paper, we look at:
% %
% \begin{inparaenum}[(A)]
%   \item \textbf{mix} of operations performed by the expressions passed to eval,
%   \item \textbf{}
% \end{inparaenum}

@@

%% Qualitatively, we observe all kinds of operations being executed inside
%% \eval calls, from numerical computing for statistical modeling to reflection
%% and metaprogramming for implementing DSLs. Even testing frameworks such
%% \c{genthat} heavily rely upon \eval to execute test code snippets. The core
%% R implementation is rife with uses of \eval for mundane things like
%% accessing default values of formal parameters and in statistical packages
%% for computation.

@@

\begin{table}[htbp] \resizebox{\columnwidth}{!}{%
    \centering
    \begin{tabular}{c|c|c|c}
        \hline
        \multicolumn{2}{c|}{Core} & \multicolumn{2}{|c}{Packages}  \\
        \hline
        Environment Class & Eval Calls \% & Environment Class & Eval Calls \% \\
        \hline
        \CoreEnvClassA & \CoreProportionA & \PackagesEnvClassA &  \PackagesProportionA\\
        \CoreEnvClassB & \CoreProportionB & \PackagesEnvClassB &  \PackagesProportionB\\
        \CoreEnvClassC & \CoreProportionC & \PackagesEnvClassC &  \PackagesProportionC\\
        \CoreEnvClassD & \CoreProportionD & \PackagesEnvClassD &  \PackagesProportionD\\
        \CoreEnvClassE & \CoreProportionE & \PackagesEnvClassE &  \PackagesProportionE\\
        \hline
    \end{tabular}} \label{tab:environments}
\caption{Environments in terms of eval calls.}
\end{table}

@@

>>>>

@@

\newcommand{\mypara}[1]{\medskip\noindent\emph{#1}\xspace}

%%%%%
\section{Usage Metrics}

This section presents a high-level picture of the usage of \eval in R.

(CHECK IF TRUE) As none of the Kaggle programs invoke \eval, the
remainder of the paper focuses solely on the CRAN data. We will return
to Kaggle in our conclusion.

We recorded \AllAllCallCountRnd calls to \eval and its variants (XXX001
for \eval, XXX002 for \c{evalq}, XXX003 for \c{eval.parent} and XXX004 for
\c{local}).

XXX005 packages in our corpus call \eval at least once.

\mypara{Number of eval call sites.}
We observed that the number of \eval call sites is small.
Figure XXX006 shows a violin plot of the distribution of call sites to
\eval and its three variants. For packages that have call sites, the
median is XXX007 call sites. \OneCallSitesEnrich{} packages have a
single call site, and one package, \emph{metafor}, has
\MaxCallSitesEnrich{} call sites. \NoCallSitesEnrich{} packages have
no calls to \eval. These numbers are under-approximations as \eval may
be aliased, called reflectively and given as argument to a
higher-order function such as \c{map}.

\newcommand{\NOTE}[1]{{\it Note: #1}\xspace}

\mypara{Number of calls to eval.} We observed an average XXX010 calls
to \eval per run. \NOTE{this is in packages that had any call to eval.
  How do we compute this? Say we are interested in package P. Do we
  only count runs that call eval from package P. Or do we also count
  runs that did not have an eval but executed some code from P? The
  latter may be easier. Which one is right?} Figure XXX011 shows a
violin plot of the distribution of calls to eval per run. The largest
number of calls is XXX012 and comes from the package XXX013.

\mypara{Amount of code loaded by eval.} We use \c{deparse} to turn the
input to \eval into strings and measure the size of that input. The
largest input is XXX008 and the median input size is XXX009. Most
inputs are small, less than XXX014 bytes. The average size is XXX015
bytes.

\mypara{Amount of computation via eval.}
We compare the number of events intercepted during execution of \eval
across the whole corpus. For this, we use the number of interpreter
steps as a proxy for events. We observe \AllEventCountRnd events all
told, and \EvalEventCountRnd events occur inside eval
(\EvalEventAllPerc of all events).
Figure~XXX015 show a violin plot of events.

@@

\mypara{The case of CORE libraries} ... \NOTE{Give numbers for the
  same things as above but for the base/core libraries} Compare.
\NOTE{Add XXX016 to XXX025.}

\section{A Taxonomy of Eval}

@@

%% Qualitatively, we observe all kinds of operations being executed
%% inside \eval calls, from numerical computing for statistical
%% modeling to reflection and metaprogramming for implementing DSLs.
%% Even testing frameworks such \c{genthat} heavily rely upon \eval to
%% execute test code snippets. The core R implementation is rife with
%% uses of \eval for mundane things like accessing default values of
%% formal parameters and in statistical packages for computation.

@@

\begin{table}[htbp]{%} \resizebox{\columnwidth}{!}{%
    \centering
    \begin{tabular}{c|c|c|c}
        \hline
        \multicolumn{2}{c|}{Core} & \multicolumn{2}{|c}{Packages}  \\
        \hline
        Environment Class & Eval Calls \% & Environment Class & Eval Calls \% \\
        \hline
        \CoreEnvClassA & \CoreProportionA & \PackagesEnvClassA &  \PackagesProportionA\\
        \CoreEnvClassB & \CoreProportionB & \PackagesEnvClassB &  \PackagesProportionB\\
        \CoreEnvClassC & \CoreProportionC & \PackagesEnvClassC &  \PackagesProportionC\\
        \CoreEnvClassD & \CoreProportionD & \PackagesEnvClassD &  \PackagesProportionD\\
        \CoreEnvClassE & \CoreProportionE & \PackagesEnvClassE &  \PackagesProportionE\\
        \hline
    \end{tabular}} \label{tab:environments}
\caption{Environments in terms of eval calls.}
\end{table}

@@

\section{Case Studies}

\section{Why do we Eval}

\section{Analysis Results}
